---
title: "CodeBook for run_analysis.R"
author: "James Chen"
date: "Friday, July 25, 2014"
output: html_document
---

Due to campatibility issue this CodeBook.Rmd can't be run with Knit HTML. 
However, it can be served as a documentation for describing the variables, 
the data and any transformation that were perfomed to clean up the data.

Requirement:
   
Create one R script called run_analysis.R that does the following. 
 
Merges the training and the test sets to create one data set.
Extracts only the measurements on the mean and standard deviation for each measurement. 
Uses descriptive activity names to name the activities in the data set
Appropriately labels the data set with descriptive variable names. 
Creates a second, independent tidy data set with the average of each variable for each activity and each subject. 


Setting Directory for Downlaoding and Unzip file
```{r}
ProjectDirectory = getwd()
DataDirectory = "UCI HAR Dataset/"
dataFile = "dataset.RData"
download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip", 
               "data.zip", quiet = TRUE, mode = "wb")  
unzip("data.zip")
```

Reading labels
```{r}
  setwd(DataDirectory)
  
  temp = read.table("activity_labels.txt", sep = "") 
  activityLabels = as.character(temp$V2)
```

Read feature names
```{r}
  temp = read.table("features.txt", sep = "")
  attributeNames = temp$V2
```
  
Read X_train data
```{r}
  Xtrain = read.table("train/X_train.txt", sep = "")
```

Setting attribute names
```{r}
names(Xtrain) = attributeNames

```
  
Read y_train data
```{r}
  Ytrain = read.table("train/y_train.txt", sep = "")
```

Settting names
```{r}
  names(Ytrain) = "Activity"
```

Factorization
```{r}
Ytrain$Activity = as.factor(Ytrain$Activity)
```

Requirement: Uses descriptive activity names to name the activities in the data set
Provides access to the levels attribute of a variable

```{r}
  levels(Ytrain$Activity) = activityLabels
```
  
Read Subject data
```{r}
  trainSubjects = read.table("train/subject_train.txt", sep = "")
  names(trainSubjects) = "subject"
```

Setting Subject as factor
```{r}
  trainSubjects$subject = as.factor(trainSubjects$subject)
```

Combing data
```{r}
  train = cbind(Xtrain, trainSubjects, Ytrain)
```

Read test data same as train data

```{r}

  Xtest = read.table("test/X_test.txt", sep = "")
  names(Xtest) = attributeNames

  Ytest = read.table("test/y_test.txt", sep = "")
  names(Ytest) = "Activity"

  Ytest$Activity = as.factor(Ytest$Activity)
  levels(Ytest$Activity) = activityLabels
  
  testSubjects = read.table("test/subject_test.txt", sep = "")
  names(testSubjects) = "subject"

  testSubjects$subject = as.factor(testSubjects$subject)
  test = cbind(Xtest, testSubjects, Ytest)
```

Save to data file
```{r}
  save(train, test, file = dataFile)
```
Remove objects for saving memoary
```{r}
  rm(train, test, temp, Ytrain, Ytest, Xtrain, Xtest, trainSubjects, testSubjects, 
     activityLabels, attributeNames)
```


Reload objects
```{r}
load(dataFile)
```

Setting working directory
```{r}
setwd(ProjectDirectory)
```

Adding new columns
```{r}
train$Partition = "Train"
test$Partition = "Test"
```
Merges the training and the test sets to create one data set.
```{r}
all = rbind(train, test)  # combine sets for visualization
all$Partition = as.factor(all$Partition)
```


Fielter then Combine
```{r}
library(plyr)
library(reshape)

cindex <-grepl(pattern = "mean\\(|std\\(", names(all), perl=TRUE)

cindex[ (length(cindex)-2):  (length(cindex)) ] <- TRUE
extract<-all[, cindex]
```

Save object fo a data file
```{r}
save(train, test, extract, file = dataFile)

```

Reload the data file
```{r}
load(dataFile)
setwd(ProjectDirectory)
```

Melt an object into a form suitable for easy casting.
```{r}
tidydata <- melt(extract,id=67:69,measure.vars=1:66)
```


Create final tidy data
```{r}
newtidydata<-ddply(tidydata,.(subject,Activity,Partition,variable),summarize,value=mean(value))
```

Final tidydata

dim(newtidydata)
[1] 11880     5

names(newtidydata)
[1] "subject"   "Activity"  "Partition" "variable"  "value"    

table(newtidydata$Activity)

           WALKING   WALKING_UPSTAIRS WALKING_DOWNSTAIRS            SITTING           STANDING             LAYING 
              1980               1980               1980               1980               1980               1980 
              
table(newtidydata$Partition)

 Test Train 
 3564  8316 

table(newtidydata$variable)


          tBodyAcc-mean()-X           tBodyAcc-mean()-Y           tBodyAcc-mean()-Z            tBodyAcc-std()-X 
                        180                         180                         180                         180 
           tBodyAcc-std()-Y            tBodyAcc-std()-Z        tGravityAcc-mean()-X        tGravityAcc-mean()-Y 
                        180                         180                         180                         180 
       tGravityAcc-mean()-Z         tGravityAcc-std()-X         tGravityAcc-std()-Y         tGravityAcc-std()-Z 
                        180                         180                         180                         180 
      tBodyAccJerk-mean()-X       tBodyAccJerk-mean()-Y       tBodyAccJerk-mean()-Z        tBodyAccJerk-std()-X 
                        180                         180                         180                         180 
       tBodyAccJerk-std()-Y        tBodyAccJerk-std()-Z          tBodyGyro-mean()-X          tBodyGyro-mean()-Y 
                        180                         180                         180                         180 
         tBodyGyro-mean()-Z           tBodyGyro-std()-X           tBodyGyro-std()-Y           tBodyGyro-std()-Z 
                        180                         180                         180                         180 
     tBodyGyroJerk-mean()-X      tBodyGyroJerk-mean()-Y      tBodyGyroJerk-mean()-Z       tBodyGyroJerk-std()-X 
                        180                         180                         180                         180 
      tBodyGyroJerk-std()-Y       tBodyGyroJerk-std()-Z          tBodyAccMag-mean()           tBodyAccMag-std() 
                        180                         180                         180                         180 
      tGravityAccMag-mean()        tGravityAccMag-std()      tBodyAccJerkMag-mean()       tBodyAccJerkMag-std() 
                        180                         180                         180                         180 
        tBodyGyroMag-mean()          tBodyGyroMag-std()     tBodyGyroJerkMag-mean()      tBodyGyroJerkMag-std() 
                        180                         180                         180                         180 
          fBodyAcc-mean()-X           fBodyAcc-mean()-Y           fBodyAcc-mean()-Z            fBodyAcc-std()-X 
                        180                         180                         180                         180 
           fBodyAcc-std()-Y            fBodyAcc-std()-Z       fBodyAccJerk-mean()-X       fBodyAccJerk-mean()-Y 
                        180                         180                         180                         180 
      fBodyAccJerk-mean()-Z        fBodyAccJerk-std()-X        fBodyAccJerk-std()-Y        fBodyAccJerk-std()-Z 
                        180                         180                         180                         180 
         fBodyGyro-mean()-X          fBodyGyro-mean()-Y          fBodyGyro-mean()-Z           fBodyGyro-std()-X 
                        180                         180                         180                         180 
          fBodyGyro-std()-Y           fBodyGyro-std()-Z          fBodyAccMag-mean()           fBodyAccMag-std() 
                        180                         180                         180                         180 
 fBodyBodyAccJerkMag-mean()   fBodyBodyAccJerkMag-std()     fBodyBodyGyroMag-mean()      fBodyBodyGyroMag-std() 
                        180                         180                         180                         180 
fBodyBodyGyroJerkMag-mean()  fBodyBodyGyroJerkMag-std() 
                        180                         180 
 
 
              

Write to a file
```{r}
write.table(newtidydata, "mytidydata.txt", sep="\t") 
```

